---
title: "Homework 10"
author: "Bailey Kretzler"
date: "4/14/2021"
output: 
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

## Question 1:

```{r}

#----------------------------------------------------------
#FUNCTION: counter
#Description: A function that counts the number of zeros in a vector
#Inputs: a numeric vector
#Outputs: the number of zeros
#----------------------------------------------------------
counter <- function(x = 0:12) {
  
  count = 0
  
  for( i in seq_along(x)){
    if(x[i] == 0){
      count = count +1
    }
  }

  #function body
  
  return(count)
  
} #end of counter
#----------------------------------------------------------

counter()

testVec <- c(0:3, -1:1, -5:8)

counter(x = testVec)

```



## Question 2:

```{r}
testVec <- c(0:3, -1:1, -5:8)


length(which(testVec == 0))



```

## Question 3:


```{r}

#----------------------------------------------------------
#FUNCTION: matmulti
#Description: creating a matrix and filling it with the row # * col #
#Inputs: Matrix dimensions
#Outputs: a filled matrix
#----------------------------------------------------------
matmulti <- function(r = 3, c = 3) {

  mat <- matrix(data = 0, nrow = r, ncol = c)
  for( i in 1:r){
    for(j in 1:c){
      mat[i,j] = i*j
    }
  }
  
  return(mat)
  
} #end of matmulti
#----------------------------------------------------------

matmulti()


```




## Question 4:


```{r}
#---------------------------------------------------------#
#                      prelims
#---------------------------------------------------------#

library(ggplot2)
library(TeachingDemos)
library(agricolae)



char2seed("croissant with almonds")



#---------------------------------------------------------#
#                      building functions
#---------------------------------------------------------#


#----------------------------------------------------------
#FUNCTION: readDat
#Description: Read in or generate data set 
#Inputs: file name or nothing (as here)
#Outputs: 3 column data fram of observed data
#----------------------------------------------------------
readDat <- function(z = NULL) {
  if(is.null((z))){
      xVal <- 1:20
      yVal <- xVal +10*rnorm(20)
      df <- data.frame(ID = seq_along(xVal),
                       x = xVal,
                       y = yVal)
  } else
  
  df <- read.table( file = z,
                    header = TRUE,
                    stringsAsFactors = FALSE,
                    sep = ",") # add seperator to indicate that this is comma seperated file
  
  return(df)
  
} #end of readDat
#----------------------------------------------------------

#readDat()


#----------------------------------------------------------
#FUNCTION: getMet
#Description: calculate metric for randomization test
#Inputs: 2 column data frame for regression, column ID for parameters to analyze
#Outputs: regression slope
#----------------------------------------------------------
getMet <- function(z = NULL, P1 = 3, P2 = 2) {
  if(is.null(z)){
        xVal <- c(rep("control",10), rep("treatment",10))
        yVal <- rnorm(20) +10*rnorm(20)
        z <- data.frame(ID = seq_along(xVal),
                          x = xVal,
                          y = yVal)
        
  } 
    
  . <- aov(z[,P1]~z[,P2]) # use anova to get p
  . <- summary(.)
  met <- .[[1]][1,5] #set metric = to the p
  
 
  
  return(met)
  
} #end of getMet
#----------------------------------------------------------

#a =summary(aov(df$Gs ~df$Water))

#getMet()


#----------------------------------------------------------
#FUNCTION: shuffleDat
#Description: randomize data for a regression
#Inputs: 3 column data frame (ID, x, y)
#Outputs: 3 column data frame (ID, x, y) - reshuffled
#----------------------------------------------------------
shuffleDat <- function(z = NULL, P1 = 3, P2 = 2) {
            if(is.null(z)){
                  xVal <- c(rep("control",10), rep("treatment",10)) # change x to a categorical
                  yVal <- rnorm(20) +10*rnorm(20)
                  z <- data.frame(ID = seq_along(xVal),
                                  x = xVal,
                                  y = yVal)} 
  
    z[,P1] <- sample(z[,P1]) # change to reshuffle specified column
 
  
  return(z)
  
} #end of shuffleDat
#----------------------------------------------------------

#shuffleDat()

#----------------------------------------------------------
#FUNCTION: getP
#Description: obtain p value from simulation
#Inputs: list of observed metric and vector of simulated metrics
#Outputs: upper and lower tail prob value
#----------------------------------------------------------
getP <- function(z = NULL) {
  if(is.null(z)){
    z <- list(rnorm(1), rnorm(1000))}
    
  plower <- mean(z[[2]] <= z[[1]]) # calculated the proportion when simulated value is less than or equal to observed value
  pupper <-  mean(z[[2]] >= z[[1]]) # calculated the prop when sim is greater than or equal to obs
  
  
  
  return(c(pL = plower, pU = pupper))
  
} #end of getP
#----------------------------------------------------------

#getP()

#----------------------------------------------------------
#FUNCTION: plotRan  
#Description: create ggplot of histogram of simulated values  
#Inputs: liat of observed metric and vector of simulated metrics
#Outputs: saved ggplot histogram
#----------------------------------------------------------
plotRan  <- function(z = NULL) {
    if(is.null(z)){
      z<- list(rnorm(1), rnorm(1000))
    }
  df <- data.frame(ID = seq_along(z[[2]]), simX = z[[2]])
  p1 <- ggplot(data = df, mapping = aes(x=simX))+ 
    geom_histogram(mapping = aes(fill = I("goldenrod"),
                                 colour = I("black")))+ 
    geom_vline(aes(xintercept = z[[1]], col = "blue"))
  print(p1)
  
} #end of plotRan 
#----------------------------------------------------------

#plotRan()


###########################################################

nSim <- 1000 # #of data sets to simulate
xSim <-  rep(NA,nSim)#set up empty vector for simdata
df <- readDat(z = "GA and KR_paper.csv") #load in data from my MSc

colnames(df) # checking the column names to see what we have

#lets pull out the genotype (GT), Water regiment (WT), Abscisic acid concentrations (ABA) and stomatal conductance (gs)

df <- df[,c("GT","Water", "ABA", "Gs")]
head(df)


xObs <- getMet(df, P1 = 4, P2 = 1) #get mean of data
print(xObs)

for(i in seq_len(nSim)){ # loop this 1000 times
  xSim[i] <- getMet(shuffleDat(df, P1 = 4), P1 = 4, P2 = 1) #create a metric for a shuffled data set
}

pvals <- list(xObs,xSim) # create a list of the two object we created

getP(pvals) # get p values
plotRan(pvals) #plot histogram


```



## Question 5



Basically what I did in question 4 was testing how different the p values are for all the simulations. The actual p value and the mean predicted P values are honestly pretty close which I find surprising. There was no difference here in stomatal conductance between genotypes. I originally tried with watering regiment but the significance was way to high to be a reasonable outcome. It resulted in a P upper of 1 because all of the simulated P values were always greater. This was due to the nature of the measurement. Stomatal conductance drops off rapidly under severe drought stress, which these wheat plants were certainly exposed to.

```{r}

# get regular p value from usual analysis

p <- summary(aov(df$Gs~df$GT))[[1]][1,5]

#versus

p2 <- getP(pvals) # predicted p values


print(c(p, p2)) # viewing them together






```

