---
title: "Homework 7"
author: "Bailey Kretzler"
date: "3/17/2021"
output: 
  prettydoc::html_pretty:
    theme: cayman
    highlight: github

---

## Simulated Data

#### Open Libraries

```{r}

library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation


```

#### Read in data vector 


###### Simulated Data

```{r}
# quick and dirty, a truncated normal distribution to work on the solution set

z <- rnorm(n=3000,mean=0.2)
z <- data.frame(1:3000,z)
names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
summary(z$myVar)

```


#### Plot Histogram of data

```{r}
p1 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)

```



#### Add empirical density curve

```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

```




#### Get maximum likelihood parameters

```{r}
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute

```



#### Plot normal probability density 

```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
 p1 + stat

```


#### Plot exponential probability density


```{r}
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$myVar), args = list(rate=rateML))
 p1 + stat + stat2

```



#### Plot uniform probability density

```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
 p1 + stat + stat2 + stat3

```



#### Plot gamma probability density

```{r}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4

```



#### Plot beta probability density

```{r}
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial

```



## My data


#### Read in data vector 


###### Intercropping data

```{r}
# loading in some data from an experiment I did in Summer 2020

z <- read.table("Rsheet_IC.csv",header=TRUE,sep=",", stringsAsFactors=FALSE)
str(z)
summary(z)


```


#### Plot Histogram of data

```{r}
#x = pounds of phosphorus per acre 
p1 <- ggplot(data=z, aes(x=P_lbA, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)

```



#### Add empirical density curve

```{r}

#add the density curve as a dotted line
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

```




#### Get maximum likelihood parameters

```{r}
#use fitdistr to fit normal distribution parameters to data
normPars <- fitdistr(z$P_lbA,"normal")

#pring to view what it generated
print(normPars)
#view the structure
str(normPars)
#extract the mean
normPars$estimate["mean"] # note structure of getting a named attribute

```



#### Plot normal probability density 

```{r}
#create mean and standard deviation values from fitdistr above
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

#creating a vector for x values to use on the graph
#zero to the max value of phoshorus
#length is the number of observations
xval <- seq(0,max(z$P_lbA),len=length(z$P_lbA))

#add a normal shaped line to the graph in red
 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$P_lbA), args = list(mean = meanML, sd = sdML))
 p1 + stat

```


#### Plot exponential probability density


```{r}
#fitdistr for exponential dis
expoPars <- fitdistr(z$P_lbA,"exponential")
#extracting the rate from the estimates generated
#by fitdistr
rateML <- expoPars$estimate["rate"]

#add an exponential dist. line to the graph in blue
stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$P_lbA), args = list(rate=rateML))
 p1 + stat + stat2
 
#this does not match

```



#### Plot uniform probability density

```{r}

#create uniform probability using data min and max in dark green
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$P_lbA), args = list(min=min(z$P_lbA), max=max(z$P_lbA)))
 p1 + stat + stat2 + stat3

```



#### Plot gamma probability density

```{r}
#fitdistr for gamma distribution
gammaPars <- fitdistr(z$P_lbA,"gamma")
#extract shape parameter
shapeML <- gammaPars$estimate["shape"]
#extract rate parameter
rateML <- gammaPars$estimate["rate"]

#add a line for gamma distribution to the plot in brown
stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$P_lbA), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4

```



#### Plot beta probability density

```{r}

#creating a special graph?
#x is proportional to max value of x plus 0.1
pSpecial <- ggplot(data=z, aes(x=P_lbA/(max(P_lbA + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

#fitdistr for beta 
betaPars <- fitdistr(x=z$P_lbA/max(z$P_lbA + 0.1),start=list(shape1=1,shape2=2),"beta")
#extract shape1
shape1ML <- betaPars$estimate["shape1"]
#extract shape 2
shape2ML <- betaPars$estimate["shape2"]


statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$P_lbA), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial

```


## Simulated data from my own data



#### Simulated data

```{r}
#gamma is best fitting although 
gammaSim <- fitdistr(z$P_lbA,"gamma")

#extract mean and sd
shapeSim <- gammaSim$estimate["shape"]
rateSim <- gammaSim$estimate["rate"]

#generate random values
simSim <- rgamma(n = length(z$P_lbA), shape = shapeSim, rate = rateSim)
str(simSim)

#plot values as histogram
p1 <- ggplot(data=z, aes(x=simSim, y=..density..)) +
  geom_histogram(color="black",fill="thistle4",size=0.2) + theme_bw() +xlim (0,325)
print(p1)


#add the density curve as a dotted line
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)


```



#### Actual data


```{r}
#x = pounds of phosphorus per acre 
p1 <- ggplot(data=z, aes(x=P_lbA, y=..density..)) +
  geom_histogram(color="black",fill="salmon4",size=0.2) + xlim(0,325) +theme_bw()
print(p1)

```



## Questions:

## 1.
**How do the two histogram profiles compare? Do you think the model is doing a good job of simulating realistic data that match your original measurements? Why or why not?**

The simulated histogram seems to be more spread out, whereas the real one is clustered more around the mean. It has done an ok job of simulating but there are a lot of holes in the graph. I think if there was a greater number of values it might look better.
        
## 2. 


**If you have entered a large data frame with many columns, try running all of the code on a different variable to see how the simulation performs.**


#### Plot Histogram of data

```{r}
#x = buckwheat density 

p1 <- ggplot(data=z, aes(x=BW_density, y=..density..)) +
  geom_histogram(color="black",fill="turquoise4",size=0.2) + 
  theme_bw()
print(p1)

```



#### Add empirical density curve

```{r}

#add the density curve as a dotted line
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)

```




#### Get maximum likelihood parameters

```{r}
#use fitdistr to fit normal distribution parameters to data
normPars <- fitdistr(z$BW_density,"normal")

#pring to view what it generated
print(normPars)
#view the structure
str(normPars)
#extract the mean
normPars$estimate["mean"] # note structure of getting a named attribute

```



#### Plot normal probability density 

```{r}
#create mean and standard deviation values from fitdistr above
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

#creating a vector for x values to use on the graph
#zero to the max value of phoshorus
#length is the number of observations
xval <- seq(0,max(z$BW_density),len=length(z$BW_density))

#add a normal shaped line to the graph in red
 stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$BW_density), args = list(mean = meanML, sd = sdML))
 p1 + stat

```


#### Plot exponential probability density


```{r}
#fitdistr for exponential dis
expoPars <- fitdistr(z$BW_density,"exponential")
#extracting the rate from the estimates generated
#by fitdistr
rateML <- expoPars$estimate["rate"]

#add an exponential dist. line to the graph in blue
stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$BW_density), args = list(rate=rateML))
 p1 + stat + stat2
 
#this does not match

```



#### Plot uniform probability density

```{r}

#create uniform probability using data min and max in dark green
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$BW_density), args = list(min=min(z$P_lbA), max=max(z$BW_density)))
 p1 + stat + stat2 + stat3

```



#### Plot gamma probability density

```{r}
#fitdistr for gamma distribution
gammaPars <- fitdistr(z$BW_density,"gamma")
#extract shape parameter
shapeML <- gammaPars$estimate["shape"]
#extract rate parameter
rateML <- gammaPars$estimate["rate"]

#add a line for gamma distribution to the plot in brown
stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="brown", n = length(z$BW_density), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4

```



#### Plot beta probability density

```{r}

#creating a special graph?
#x is proportional to max value of x plus 0.1
pSpecial <- ggplot(data=z, aes(x=BW_density/(max(BW_density + 0.1)), y=..density..)) +
  geom_histogram(color="black",fill="turquoise4",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted") + theme_bw()

#fitdistr for beta 
betaPars <- fitdistr(x=z$BW_density/max(z$BW_density + 0.1),start=list(shape1=1,shape2=2),"beta")
#extract shape1
shape1ML <- betaPars$estimate["shape1"]
#extract shape 2
shape2ML <- betaPars$estimate["shape2"]


statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$BW_density), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial

```


For this variable, the gamma distribution looks to be the best fit.
